%!TEX TS-program = xelatex
\documentclass[a4paper,online]{adcv}

\usepackage[english]{babel}

\addbibresource{bib_alessandro_rossini.bib}

\title{Jyotirmoy Banerjee’s CV}

\adcvname{Jyotirmoy}{Banerjee}{Ph.D.}
\adcvtitle{AI \& ML Scientist}
\adcvaddress{B-1, Nigam Nivas}{Kaggadasapura}{Bengaluru}{India}
\adcvwebsite{https://www.linkedin.com/in/jyotirmoybanerjee/}{LinkedIn}
\adcvemail{jyoban}{gmail}{com}
\adcvphone{(+91) 9611579158}
\adcvdate{Feb 2021}

\begin{document}

Jyotirmoy Banerjee is a Senior Data Scientist at Flipkart in Bengaluru, India. He has more than 12 years of experience in artificial intelligence and machine learning.
Jyotirmoy specialises in computer vision and natural language processing with applications to healthcare and e-commerce.

In his current position, Jyotirmoy helps the organisation with data science road map, mentors team members, and contributes to various projects. 
He is actively involved in other organisational efforts like mergers and acquisitions. He actively participates in fostering academic relationship.

Jyotirmoy holds a Ph.D. in computer science from the Erasmus University Rotterdam, Netherlands. Followed by a Postdoc from the University College London, United Kingdom. He has more than 20 peer-reviewed scientific publications as well as 4 patents. He is reviewer in several conferences and journals.
Jyotirmoy is fluent in English and multiple Indian languages.

\section{Education}

\begin{adcvtabletwo}
  \adcvrowtwo{\textbf{Post Doctorate}, Computer science, University College London, UK.}{}
  \adcvrowtwo{\textbf{Ph.D.}, Computer science, Erasmus University Rotterdam, Netherlands.}{}
  \adcvrowtwo{\textbf{MS by Research}, Computer science, IIIT-H, India.}{}
\end{adcvtabletwo}


\section{Experience}

\begin{adcvtabletwo}
  \adcvrowtwo{\textbf{Senior Data Scientist}, User Acquisition and Retention, Flipkart, Bengaluru, India}{2018--Present}
  \adcvrowmulti{Contributed to data science road map. Lead several data science projects including size and fit recommendation, content moderation, visual attribute extraction, and aspect extraction from reviews. Helped in deployment of machine learning models. These models used supervised approaches like deep learning and unsupervised approaches like topic modelling. Assisted in mergers and acquisitions.  Lead academic collaboration with IIT Kharagpur on auto question answering. Actively participated in hiring. Published in reputed international conference.}

  \adcvrowtwo{\textbf{Applied Scientist}, India Machine Learning, Amazon, Bengaluru, India}{2017--2018}
  \adcvrowmulti{Lead efforts towards addressing last mile delivery issues. Developed recurrence neural network based models for business hours prediction. Developed low latency machine learning models for online use cases. Supported successful transition of machine leaning models to product.}

  \adcvrowtwo{\textbf{Postdoctoral Research Associate}, Department of Medical Physics and Bioengineering, University College London, London, UK}{2016--2017}
  \adcvrowmulti{Developed advanced computer vision techniques for congenial problems in the womb, including spina bifida and twin twin syndrome. Improving visualization using non-linear optimisation techniques. Published in reputed international conference.}

  \adcvrowtwo{\textbf{Scientific Researcher}, Departments of Radiology \& Nuclear Medicine and Medical Informatics, Erasmus Medical Center, Rotterdam, Netherlands}{2011--2015}
  \adcvrowmulti{Developed method to improve image guidance in minimally invasive treatment of liver for transjugular intrahepatic portosystemic shunt treatment. Image guidance in RF ablation procedures, using image fusion of CT and ultrasound. The real-time 4D ultrasound tracking technology was transferred to Philips. Published in reputed international conferences and journals.}
  
  \adcvrowtwo{\textbf{Scientist}, Biomedical Imaging Group, GE Global Research, Bengaluru, India}{2009--2011}
  \adcvrowmulti{Developed image guidance for catheter insertion in neonatal. Automated the process of identifying the fetal head for obstetric application. Published in reputed international conference. Contributed to key IP for the program and helped transition to product.}

  \adcvrowtwo{\textbf{Research Assistant}, Centre for Visual Information Technology, IIIT-H, Hyderabad, India}{2005--2008}
  \adcvrowmulti{Worked on document enhancement projects. Published in reputed conference. Received best paper award.}
  
  \adcvrowtwo{\textbf{Software Engineer}, Perot Systems, Bengaluru, India}{2003--2005}
  \adcvrowmulti{Contributed to healthcare payer application software. Incorporated specific business requirements for global clients.}

%  \fi
%  \adcvrowskip
%  \adcvrowtwo{\textbf{Lecturer}, Dep. of Electrical and Electronic Eng., \hib{}, Norway}{2012}
%  \ifextended
%    \adcvrowmulti{Teaching of the Bachelor course Industrial Software Engineering (see \hyperref[sec:teaching]{Teaching} below).}
%  \fi
\end{adcvtabletwo}

\section{Patents and Publications}

\href{https://scholar.google.co.in/citations?user=SaF-0PcAAAAJ&hl=en}{4 patents and 20+ publications} in peer-reviewed conference proceedings and journals.

\section{Awards}

Best paper award for the paper - `Super-resolution of Text Images Using Edge-Directed Tangent Field'.

\section{Reviewer}

Served as reviewer in multiple international conferences and journals, namely, IEEE Signal Processing, Elsevier Digital Signal Processing, Elsevier Pattern Recognition Letters, IEEE TMI, ECCV 2014 workshop, ICVGIP 2014 \& 2016.

\section{Skills}

\textbf{Primary}: Python, Pytorch, Tensorflow, OpenCV, C++.

\textbf{Secondary}: Machine Learning, Computer Vision, Natural Language Processing, Deep learning.


\section{Personal interests}

Sports, Science \& Technology, Traveling, Movies, Music.

\end{document}
